Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-07-12 prop_letter.tex
prop_letter.tex After merge small changes to main text...

The question now is what happens if we relax the perfect measurement
assumption that the intervals are precisely $[0,0]$ and
$[1,1]$. We need to examine what happens if
measurement imperfections occasionally cause an
ideally impossible event to be detected or an ideally certain event to
be not be detected. Mermin argues in this case that:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}

We can formally test this hypothesis by considering the notion of
$\delta$-deterministic QIVPM below.

\begin{definition}\label{def:delta-deterministic}A QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$
  is called $\delta$-deterministic if either
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$ for any
  $P\in\events$.\end{definition}

A cryptodeterministic QIVPM is $0$-deterministic.  As $\delta$ gets
bigger, the QIVPM allows more and more indeterminate behavior. A
$\frac{1}{3}$-deterministic QIVPM can be given by
$\bar{\mu}\left(P\right)=\left[x/3,x/3\right]$, where $P$ projects
onto a $x$-dimensional subspace in a three-dimensional Hilbert space.

We now prove that for small enough $\delta$, the Kochen-Specker is
still valid, i.e., if the measurement imperfections are small enough,
we cannot nullify the Kochen-Specker theorem as
Meyer~\cite{PhysRevLett.83.3751} argued.

@@

It remains to verify $\bar{\mu}^{\textrm{D}}$ is a cryptodeterministic
QIVPM. The main part of that verification is to check:
\begin{subequations}
\begin{eqnarray}
 &  & \bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs\\
 & \Rightarrow & \bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs
\end{eqnarray}
\end{subequations}for orthogonal operators\nb$P_{0}$ and $P_{1}$.

When $\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and $\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$
which implies $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
Since $\delta<\frac{1}{3}\Rightarrow2\delta<1-\delta$, $\bar{\mu}\left(P_{0}+P_{1}\right)$
cannot be a subset of $\left[1-\delta,1\right]$, but $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or $\left[1-\delta,1\right]$.
Hence, $\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of $\left[0,\delta\right]$,
which implies $\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$. 
\end{proof}

@@

In the idealized situation of real-valued quantum probability
measures, Gleason's theorem asserts that every measure is determined
by a unique mixed state. The theorem above is weaker however: it
asserts that, for each classical sub-event subspace, a QIVPM must be
consistent with \emph{some} state. But the states consistent with
different subspaces may be incomparable as the following example
shows.

\begin{example}\label{ex:three-dimensional-three-value}
  Given a three dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.  Let
  $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
  $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}_{3}(\mathbb{0})=\bar{\mu}_{3}(\proj{0})\\
 & =\bar{\mu}_{3}(\proj{\ps})=\bar{\mu}_{3}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}_{3}(\mathbb{1})=\bar{\mu}_{3}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}_{3}(\mathbb{1}-\proj{\ps})=\bar{\mu}_{3}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}_{3}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}_{3}$ is a QIVPM. Now
let $S_{1}=\left\{ \proj{0},\proj{\ps}\right\} $. Because
$\bar{\mu}_{3}(\proj{0})=\bar{\mu}_{3}(\proj{\ps})=\imposs$, we have
$\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=0$
for any mixed
state\nb$\rho\in\coreBorn\left(\bar{\mu}_{3},S_{1}\right)$.  Hence,
$\rho$ cannot be any state other than $\ket{2}$, the only pure state
orthogonal to both $\ket{0}$ and $\ket{\ps}$, and it is easy to verify
$\proj{2}$ is actually in
$\coreBorn\left(\bar{\mu}_{3},S_{1}\right)$. Similarly, let
$S_{2}=\left\{ \proj{0},\proj{\ps'}\right\} $. We have
$\coreBorn\left(\bar{\mu}_{3},S_{2}\right)=\left\{ \proj{1}\right\} $.
Therefore,
\begin{equation}
\coreBorn\left(\bar{\mu}_{3},\events\right)\subseteq\coreBorn\left(\bar{\mu}_{3},S_{1}\right)\cap\coreBorn\left(\bar{\mu}_{3},S_{2}\right)=\emptyset
\end{equation}
Thus, there is no state that is consistent with the QIVPM on all
projectors\nb$\events$ invalidating Gleason's theorem.
\end{example}

@@

We call a state $\rho$ a \emph{totally mixed state} if
$0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$ for any
$P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $. We also
define a notion of distance based on the following norm of a
matrix\nb$A$ as
$\left\Vert A\right\Vert
=\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$\nb\citep{544199}. Now
consider the following collection of intervals:
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\left\{ \left[0,0\right],\left[1,1\right]\right\} 
\end{eqnarray}
which are overlapping intervals with length\nb$1/n$ except the intervals
near the ends.

A totally mixed state induces a QIVPM over the intervals above as
follows. We first define a map $\widehat{\iota}_{n}$ from the interval
$[0,1]$ to the collection of intervals above as follows:
\begin{itemize}
\item when $x=0$ or $x=1$, $\widehat{\iota}_{n}\left(x\right)=\left[x,x\right]$;
\item when $0<x<1$, $\widehat{\iota}_{n}\left(x\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$,
where $k$ is defined by one of the following:
\begin{itemize}
\item $2\mid k$ and $k-0.5\le2nx\le k+0.5$;
\item $2\nmid k$ and $k-0.5<2nx<k+0.5$.
\end{itemize}
\end{itemize}

It is now not difficult to verify that for any totally mixed
state\nb$\rho$ on a three-dimensional Hilbert space,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is a
QIVPM. Furthemore, we have:
\begin{itemize}
\item If $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$,
$\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
\item If $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.
\end{itemize}
In other words, the states consistent with this QIVMP
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ will be closer and
closer to $\rho$ as $n$ goes to infinity. In the limiting case when
the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as an real-valued one and will have a unique state consistent
with it on every projector when $d\ge3$.

>>>>

The question now is what happens if we relax the perfect measurement
assumption that the intervals are precisely $[0,0]$ and
$[1,1]$. We need to examine what happens if
measurement imperfections occasionally cause an
ideally impossible event to be detected or an ideally certain event
not to be detected. Mermin argues in this case that:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}

We can formally test this hypothesis by considering the notion of
a $\delta$-deterministic QIVPM below.

\begin{definition}\label{def:delta-deterministic}We define
  QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$
  as $\delta$-deterministic if either
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$ for any
  $P\in\events$.\end{definition}

A cryptodeterministic QIVPM is $0$-deterministic.  As $\delta$ gets
bigger, the QIVPM allows more and more indeterminate behavior. A
$\frac{1}{3}$-deterministic QIVPM can be given by
$\bar{\mu}\left(P\right)=\left[x/3,x/3\right]$, where $P$ projects
onto an $x$-dimensional subspace of a three-dimensional Hilbert space.

We now prove that for small enough $\delta$, the Kochen-Specker result is
still valid, i.e., if the measurement imperfections are small enough,
we cannot nullify the Kochen-Specker theorem as
Meyer~\cite{PhysRevLett.83.3751} argued.

@@

It remains to verify that $\bar{\mu}^{\textrm{D}}$ is a cryptodeterministic
QIVPM by checking the properties of
\begin{subequations}
\begin{eqnarray}
 &  & \bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs\\
 & \Rightarrow & \bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs
\end{eqnarray}
\end{subequations}for orthogonal operators\nb$P_{0}$ and $P_{1}$.

When $\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and $\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$
which implies $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
Since $\delta<\frac{1}{3}\Rightarrow2\delta<1-\delta$, $\bar{\mu}\left(P_{0}+P_{1}\right)$
cannot be a subset of $\left[1-\delta,1\right]$, but $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or $\left[1-\delta,1\right]$.
Hence, $\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of $\left[0,\delta\right]$,
which implies $\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$. Thus the
theorem holds when $\delta<\frac{1}{3}$.
\end{proof}

@@

In the idealized situation of real-valued quantum probability
measures, Gleason's theorem asserts that every measure is determined
by a unique mixed state. The theorem above is weaker: it
asserts that, for each classical sub-event subspace, a QIVPM must be
consistent with \emph{some} state. But the states consistent with
different subspaces may be incomparable, as the following example
shows.

\begin{example}\label{ex:three-dimensional-three-value}
  Given a three-dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,  let
  $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
  $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}_{3}(\mathbb{0})=\bar{\mu}_{3}(\proj{0})\\
 & =\bar{\mu}_{3}(\proj{\ps})=\bar{\mu}_{3}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}_{3}(\mathbb{1})=\bar{\mu}_{3}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}_{3}(\mathbb{1}-\proj{\ps})=\bar{\mu}_{3}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}_{3}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}_{3}$ is a QIVPM. Now
let $S_{1}=\left\{ \proj{0},\proj{\ps}\right\} $. Because
$\bar{\mu}_{3}(\proj{0})=\bar{\mu}_{3}(\proj{\ps})=\imposs$, we have
$\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=0$
for any mixed
state\nb$\rho\in\coreBorn\left(\bar{\mu}_{3},S_{1}\right)$.  Hence,
$\rho$ cannot be any state other than $\ket{2}$, the only pure state
orthogonal to both $\ket{0}$ and $\ket{\ps}$, and it is easy to verify
$\proj{2}$ is actually in
$\coreBorn\left(\bar{\mu}_{3},S_{1}\right)$. Similarly, let
$S_{2}=\left\{ \proj{0},\proj{\ps'}\right\} $. We have
$\coreBorn\left(\bar{\mu}_{3},S_{2}\right)=\left\{ \proj{1}\right\} $.
Therefore,
\begin{equation}
\coreBorn\left(\bar{\mu}_{3},\events\right)\subseteq\coreBorn\left(\bar{\mu}_{3},S_{1}\right)\cap\coreBorn\left(\bar{\mu}_{3},S_{2}\right)=\emptyset
\end{equation}
Thus, there is no state that is consistent with the QIVPM on all
projectors\nb$\events$, so Gleason's theorem does not hold for at least this
example.
\end{example}

@@

We call a state $\rho$ a \emph{totally mixed state} if
$0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$ for any
$P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $. We also
define a notion of distance based on the following norm of a
matrix\nb$A$ as
$\left\Vert A\right\Vert
=\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{A}{\psi}\right|$\nb\citep{544199}. Now
consider the following collection of intervals:
\begin{eqnarray}
\widehat{\mathscr{I}}_{n} & = & \set{\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\nonumber \\
 &  & \cup\left\{ \left[0,0\right],\left[1,1\right]\right\} 
\end{eqnarray}
which are overlapping intervals with length\nb$1/n$ except the intervals
at each end.

A totally mixed state induces a QIVPM over the intervals above as
follows. We first define a map $\widehat{\iota}_{n}$ from the interval
$[0,1]$ to the collection of intervals as:
\begin{itemize}
\item when $x=0$ or $x=1$, $\widehat{\iota}_{n}\left(x\right)=\left[x,x\right]$;
\item when $0<x<1$, $\widehat{\iota}_{n}\left(x\right)=\left[\frac{k-1}{2n},\frac{k+1}{2n}\right]\cap\left[0,1\right]$,
where $k$ is defined by one of the following:
\begin{enumerate}
\item $k$ even and $k-0.5\le2nx\le k+0.5$;
\item $k$ odd and $k-0.5<2nx<k+0.5$.
\end{enumerate}
\end{itemize}

It is now not difficult to verify that for any totally mixed
state\nb$\rho$ on a three-dimensional Hilbert space,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is a
QIVPM. Furthemore, we have:
\begin{itemize}
\item If $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$,
$\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
\item If $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.
\end{itemize}
In other words, for the states consistent with this QIVPM,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ will become closer and
closer to $\rho$ as $n$ goes to infinity. In the limiting case when
the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as a real-valued one and will have a unique state consistent
with it on every projector when $d\ge3$.
